\section{Reproducibility}\label{sec:reproducibility}

This paper aims to score 2 on the Reproducibility Disclosure Score rubric proposed in \citet{Khan2026Survey}: \emph{(i) code public}, \emph{(ii) data openly accessible}.

\textit{Code.} The \texttt{vol-eval} package is at \url{https://github.com/ayk5511/vol-eval}, MIT-licensed. The seven-model demonstration of \Cref{sec:demonstration} is produced by \texttt{studies/01\_paper2\_demonstration.py} in the same repository, which loads the upstream Paper 2 forecast panel and writes every numerical value reported in \Cref{sec:demonstration} to JSON files in \texttt{studies/results/}. The Phase 2 reproductions of \Cref{sec:phase2} are produced by 14 per-paper scripts at \texttt{studies/per\_paper/<paper\_id>/reproduce.py}, plus three pipeline scripts (\texttt{studies/04\_search\_candidates.py}, \texttt{studies/05\_narrow\_candidates.py}, and \texttt{studies/06\_check\_unpaywall.py}) that document how the 30-paper sample was constructed. Each reproduction directory contains a \texttt{reproduction\_log.json} that records the data source, hyperparameter choices, and any approximations relative to the original paper.

\textit{Data.} The seven-model demonstration's upstream forecast panel is at \url{https://github.com/ayk5511/volatility-forecasting/blob/main/results/forecasts_5d.parquet}, CC0 licensed. It is reproducible from the upstream paper's data-collection scripts (Yahoo Finance public data) in under five minutes from a clean Python environment. The Phase 2 reproductions use Yahoo Finance, CoinMarketCap-equivalent yfinance crypto tickers, and the Borsa Istanbul / SSE Composite indices (also through yfinance); per-paper data sources are recorded in each \texttt{reproduction\_log.json}.

\textit{30-paper sample artifacts.} The full sample of 30 papers, including paper IDs, DOIs / arXiv IDs, citation counts, RDS scores (strict reading), headline-winner extractions, and per-paper voleval results, is at \texttt{studies/candidate\_30.csv}, \texttt{studies/results/phase2\_rds\_strict.csv}, \texttt{studies/results/phase2\_headline\_models.csv}, and \texttt{studies/results/phase2\_summary.json}. The source PDFs of the 30 papers are gitignored to keep the repository size manageable; the README contains instructions for re-acquisition (Crossref / arXiv / MDPI direct download).

\textit{Audit.} The repository includes \texttt{paper/audit.py}, which re-derives every numerical claim in this paper from the JSON outputs of the demonstration script and the Phase 2 aggregate. The audit script must print \texttt{ALL CHECKS PASSED} (exit 0) for the paper to be considered ship-ready. We adopted this convention after an earlier paper in this series \citep{KhanVol2026} had a sub-table whose values had been transcribed from imagination rather than from the JSON outputs; the audit-script convention catches that class of error mechanically. In the writing of the Phase 2 section, the audit script caught two transcription errors before any push.

\textit{Issues.} Corrections, extensions, and challenges to the rankings reported here are welcomed via the repository's issue tracker (\url{https://github.com/ayk5511/vol-eval/issues}). Issues become part of the public record and feed the next version of this paper. We particularly invite corrections from the original authors of any of the 30 Phase 2 papers: where our reproduction misses a hyperparameter choice that the original paper specified, where we substituted a library or feature we should not have, or where the original paper's actual data is more accessible than we believed at the time of analysis. Corrections of this kind will be reflected in version 2 of this paper.

\textit{Self-citation network.} This paper extends the methodological agenda of \citet{Khan2026Survey} (the Reproducibility Disclosure Score) and the empirical agenda of \citet{KhanVol2026} (the seven-model demonstration). It now also delivers the Phase 2 empirical extension that the v0 of this paper described as ``in production''; the reproduction work spans April 2026 and is dated to that period.
