\documentclass[12pt,a4paper]{article}

\usepackage[utf8]{inputenc}
\usepackage[T1]{fontenc}
\usepackage{lmodern}
\usepackage[margin=1in]{geometry}
\usepackage{amsmath,amssymb,amsfonts}
\usepackage{graphicx}
\usepackage{booktabs}
\usepackage{longtable}
\usepackage{array}
\usepackage{tabularx}
\usepackage{hyperref}
\usepackage{cleveref}
\usepackage{natbib}
\usepackage{setspace}
\usepackage{enumitem}
\usepackage{xcolor}
\usepackage{float}
\usepackage{listings}

\hypersetup{
    colorlinks=true,
    linkcolor=blue!70!black,
    citecolor=green!50!black,
    urlcolor=blue!70!black
}

\onehalfspacing

\lstset{
    basicstyle=\ttfamily\small,
    backgroundcolor=\color{gray!5},
    frame=single,
    framerule=0.4pt,
    breaklines=true,
    showstringspaces=false,
    columns=flexible
}

\title{vol-eval: A Python Package for Volatility\\
Forecast Evaluation, with a Re-Analysis of\\
30 Published Comparison Papers (2018--2025)}

\author{
    Akram Khan\thanks{Independent Researcher. ORCID: \href{https://orcid.org/0009-0002-7521-8648}{0009-0002-7521-8648}. Email: \href{mailto:1819ak@gmail.com}{1819ak@gmail.com}. Replication code, data, and audit script: \href{https://github.com/ayk5511/vol-eval}{github.com/ayk5511/vol-eval}.}
}

\date{April 2026}

\begin{document}

\maketitle

\begin{abstract}
This paper introduces \texttt{vol-eval}, an open-source Python package that ships the canonical battery of significance tests for volatility forecast evaluation. The battery covers QLIKE, MSE, MAE, Mincer-Zarnowitz $R^2$, Diebold-Mariano (1995) with Newey-West HAC standard errors, the Hansen-Lunde-Nason (2011) Model Confidence Set with stationary bootstrap, Hansen's (2005) Superior Predictive Ability test with three p-value variants, White's (2000) Reality Check, and the Giacomini-White (2006) Conditional Predictive Ability test. The package follows a functional API with typed result objects, ships under MIT license, and is verified by 45 unit tests including cross-validation against the \texttt{arch} package's reference implementation.

We apply \texttt{vol-eval} in two settings. The first is a self-contained seven-model horse race on S\&P 500 realized volatility, using the panel of \citet{KhanVol2026}. The second is a sample of 30 published volatility-forecasting comparison papers from 2018--2025, drawn from the \emph{Journal of Financial Econometrics}, three open-access MDPI venues, and arXiv q-fin.

Three layered findings emerge. First, a disclosure crisis. The strict-reading mean Reproducibility Disclosure Score across the 30-paper sample is 0.47, materially lower than the 0.78 reported by \citet{Khan2026Survey} for the broader 99-paper ML-finance sample. Zero of the 30 papers ship both code and publicly accessible data. Second, a methodological gap. 27 of 30 papers (90\%) declare a headline winner without applying any formal predictive-accuracy test (DM, MCS, SPA, or Reality Check). Third, a robustness gap. In the 14 papers we could reproduce (the strict-RDS-1 subset, where data is publicly accessible), 12 paper-claimed wins (86\%) do not survive significance testing under the \texttt{vol-eval} battery. Only one paper has its claim survive in full, with one additional paper showing partial survival on a subset of assets. Each per-paper reproduction and its caveats are documented in detail.

The full \texttt{vol-eval} source, the per-paper reproductions, and an audit script that re-derives every numerical claim in this paper from JSON outputs are released at \href{https://github.com/ayk5511/vol-eval}{github.com/ayk5511/vol-eval}, scoring 2 on the Reproducibility Disclosure Score rubric of \citet{Khan2026Survey}.

\medskip
\noindent\textbf{Keywords:} volatility forecasting, Diebold-Mariano test, Model Confidence Set, Hansen SPA, Reality Check, QLIKE, forecast evaluation, reproducibility, open-source software

\medskip
\noindent\textbf{JEL Classification:} C12, C15, C22, C52, C53, G17
\end{abstract}

\newpage
\tableofcontents
\newpage

\input{sections/01_introduction}
\input{sections/02_package}
\input{sections/03_methodology}
\input{sections/04_demonstration}
\input{sections/05_phase2}
\input{sections/06_discussion}
\input{sections/07_reproducibility}

\bibliographystyle{apalike}
\bibliography{bib/references}

\end{document}
